A simulation model of DNA methylation data

### 2015-04-03 Rmarkdown/methsim.Rmd
Section on simulating reads, i.e., methsim::simulate,SimulateBSParam-method.

\subsection{Simulating the truth}

@@

\subsubsection{Simulating multiple samples from the same group}

>>>>

\subsection{Simulating a single methylome}

@@

\subsection{Simulating reads}

Once we have our 'true' methylome, we want to simulate an assay of this sample. In theory this could be any type of methylation assay, but here we focus on simulating whole-genome bisulfite-sequencing data.

`methsim` uses a simple method for simulating bisulfite-sequencing. The user specifies the desired read length and average sequencing coverage[^read_length], the error rate ($\epsilon$, which includes both sequencing error and bisulfite-conversion error), and provides the simulated 'true' methylome.

The number of reads required is given by $n_{reads} = \frac{\text{average sequencing coverage}}{\text{read length}} \times \text{size of genome}$. The number of reads per chromosome is assigned proportional to the chromosome length. Then, the start of each read is sampled uniformly across the chromosome respective chromosome.

[^read_length]: Currently only single-end data are supported. This isn't a big issue. Most paired-end bisulfite-sequencing datasets have overlapping mates, and so paired-end data can be crudely simulated by simply doubling the read length. Also, all reads must have the same length, i.e., no simulation of read trimming. I don't consider this feature a priority, but it could easily be implemented by generating read lengths from a probability distribution, e.g., Poisson.

For all reads that overlap a methylation locus in the 'true' methylome, a pseudo-haplotype is sampled with probabilities given by $W_{i, \cdot}$. The methylation pattern of the sampled pseudo-haplotype is then copied, with probability $\epsilon$ that the methylation state at each locus is switched.

The result is an object storing the genomic co-ordinates and methylation patterns of every read that contains a methylation locus.
